---
title: "Non-HAF Weighed Anlaysis Template"
author: "Alex Traynor"
date: "8/1/2023"
output:
  html_document:
    df_print: paged
    output: cosmo
  pdf_document: default
---

## Setup
```{r setup, message=FALSE}

.libPaths("C:/R-4.2.1/library")

#devtools::install_git(url = "https://github.com/nstauffer/aim.analysis/", ref = "master")

### Package Setup
# Set local library path and load required packages
library(aim.analysis)
library(sp)
library(dplyr)
library(stringr)
library(sf)
library(rgdal)
library(tidyverse)
library(scales)
library(plotly)
library(arcgisbinding)
arc.check_product() 
```

### San Luis Watershed
### Set variables and input parameters

```{r input variables, message=FALSE}
#Alber's Equal Area is useful for calculating areas
projection <- sp::CRS("+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=37.5 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")
#Analysis date
date <- Sys.Date()
# Confidence level in percent
confidence <- 80
output_path <- "C:\\Users\\alaurencetraynor\\Documents\\CO\\SLVFO Restoration\\outputs"
# Filename for benchmark tool from analysis requester
benchmarks_filename <- "Monitoring Objectives.xlsx"
# Core filename for all outputs/name of reporting unit
output_filename <- "reporting_unit_name"
# The variable in AIM sample designs that contains the unique identifier for each point
aim_idvar <- "PlotID"
# The variable in AIM sample designs that contains the fate for each point
aim_fatevar = "EvalStatus"
# The variable in TerrADat that contains the unique identifier for each point matching AIM designs'
tdat_idvar <- "PrimaryKey"
# The fates in sample designs which correspond to sampled/observed
observed_fates = c("TS", "Target Sampled", "Sampled", "Eval")
# The fates in sample designs which correspond to unneeded
# (e.g. unused oversample or points designated for future sampling)
invalid_fates = c("NN", NA, "Non Needed", "", "Not Sampled")
#sdd points
design_points_name <- "CO_rejections_alamosa_trinchera_strata"
#terradat clipped points
sampled_points_name <- "Benchmarked_points_alamosa"
# The variable that contains the areas in HECTARES in wgtcats
wgtcat_area_var = "area_acres_aea"
# Reporting unit polygon/layer
wgtcats_name <- "Alamosa_trinchera_poststrata"
# Geodatabase containing all spatial files
analysis_gdb <- "Weighted Analysis Inputs.gdb"
# The unique ID for poststratification polygons
wgtcats_var <- "wgtcat"
### The filepath to the folder where we have all the spatial data
path_spatial <- "U:\\My Documents\\Analysis\\Restoration\\SLVFO Restoration\\"
### The filepath to the folder where we have the Excel stuff
path_tabular <- "U:\\My Documents\\Analysis\\Restoration\\SLVFO Restoration\\"
#my favourite function
`%notin%` <- Negate(`%in%`)
#benchmark group var in design points
benchmark_var <- "BenchmarkGroup"
# These are the fields in sampled points which contain the condition categories
benchmark_fields <-   
```

## Import Spatial and Tabular Data
### Read in weight categories
```{r read strata, message=FALSE, warning=FALSE}
#### Strata polygons
wgtcat_df <-sf::st_read(dsn = paste(path_spatial, analysis_gdb, sep = "/"),
                        layer = wgtcats_name,
                        stringsAsFactors = FALSE)
## Standardize projection
wgtcat_df <- sf::st_transform(wgtcat_df,
                              crs = projection)

## Make spatial 
wgtcat_spdf <- methods::as(wgtcat_df, "Spatial")
wgtcat_spdf@data[["wgtcat"]] <- wgtcat_spdf@data[[wgtcats_var]]
sf::st_geometry(wgtcat_df) <- NULL

wgtcat_df[["wgtcat"]] <- wgtcat_df[[wgtcats_var]]
```
### Read design points
```{r read design points, message=FALSE, warning=FALSE}
#### Read in design points
# Make sure points and polygons do not include Z/M dimensions...
design_points <- sf::st_read(dsn = paste(path_spatial, analysis_gdb, sep = "/"),
                             layer = design_points_name,
                             stringsAsFactors = FALSE)
## Standardize projection
design_points <- sf::st_transform(design_points,
                                  crs = projection)
## Make spatial 
design_points <- methods::as(design_points, "Spatial")

# Standardise columns
design_points@data[["unique_id"]] <- design_points@data[[aim_idvar]]
design_points@data[["fate"]] <- design_points@data[[aim_fatevar]]
design_points@data[["wgtcat"]] <- design_points@data[[wgtcats_var]]
design_points@data[["Benchmark.Group"]] <- NA # these are just rejections so dont have benchmark group

```
### Read benchmarks

```{r read benchmarks, message=FALSE}
#### Benchmarks
# Ive already applied these in ArcPro but they may come in handy for figures and what not
# make sure all columns (including required proportions) are filled out - otherwise apply_benchamrks() will give errors
benchmarks <- read_benchmarks(filename = benchmarks_filename,
                              filepath = path_tabular)

benchmarks$Reporting.Unit <- output_filename
```

### Read Sampled points
```{r sampled points, message=FALSE, warning=FALSE}
#### Read in Terradat points
# Make sure points and polygons do not include Z/M dimensions...
sampled_points <- sf::st_read(dsn = paste(path_spatial, analysis_gdb, sep = "/"),
                              layer = sampled_points_name,
                              stringsAsFactors = FALSE)
## Standardize projection
sampled_points <- sf::st_transform(sampled_points,
                                   crs = projection)
## Make spatial 
sampled_points <- methods::as(sampled_points, "Spatial")

sampled_points@data[["fate"]] <- "TS"

# Standardise columns
sampled_points@data[["unique_id"]] <- sampled_points@data[[tdat_idvar]]
sampled_points@data[["wgtcat"]] <- sampled_points@data[[wgtcats_var]]
sampled_points@data[["wgtcat"]] <- as.character(sampled_points@data[["wgtcat"]])
sampled_points@data$Reporting.Unit <- output_filename

```
### Filter sampled points
```{r filter sampled points, message=FALSE}
# double check and remove targeted points

# double check sampling order and design holes
 
# remove points that didnt fall in a benchmark group - well also need to double check these polys are removed from the inference area
# we can check this by looking for nulls in benchmark fields

for(field in benchmark_fields){
  sampled_points <- sampled_points[!is.na(sampled_points@data[[field]]),]
}

### Combine sampled points with rejections
# add benchmark fields to rejections
for(col in benchmark_fields){
  design_points[[col]] <- NA
}

all_points <- rbind(design_points[, c("unique_id", "fate", "wgtcat", benchmark_fields)], 
                    sampled_points[, c("unique_id", "fate", "wgtcat", benchmark_fields)])
```

## Weighting
```{r point weighting, message=FALSE, warning=FALSE}
###### Analyze ######
weight_info <- weight.gen(pts = all_points,# remove targeted plots
                          pts.fatefield = "fate",
                          pts.groupfield = "wgtcat",
                          frame.spdf = wgtcat_spdf,
                          frame.groupfield = "wgtcat",
                          target.values = c("Target Sampled", "TARGET SAMPLED", "TS"),
                          unknown.values = c("Unknown", "UNKNOWN", "UNK"),
                          nontarget.values = c("NT"),
                          inaccessible.values = c("IA"),
                          unneeded.values = invalid_fates)

point_weights <-  weight_info[["point.weights"]]

all_points <-  aim.analysis::add_coords(all_points,
                                        xynames = c("x", "y"))
```

### Organise Indicators
```{r}
# make indicators long instead of wide
benchmarked_points <- all_points@data %>% 
  pivot_longer(cols = benchmark_fields, names_to = 'Indicator', values_to = "Condition.Category")

# remove the unsampled plots
benchmarked_points <- benchmarked_points[!is.na(benchmarked_points$Condition.Category),]

#Add reporting unit
all_points[["Reporting.Unit"]] <- output_filename
weight_info[["point.weights"]][["Reporting.Unit"]] <- output_filename
# need to add reporting unit name to analyse
point_weights$Reporting.Unit <- output_filename
```

### Category Definitions
```{r prep for analyze, message=FALSE}
# a definitions list for analyze_cat_multi()
# this is just a list of all the possible condition categories
cat_defs <-  benchmarks[,c("Benchmark.Group", "Objective.name", "Condition.Category")]

colnames(cat_defs) <- c("Benchmark.Group","Indicator", "Condition.Category")
cat_defs <- unique(cat_defs)
```

## Analyse
```{r analyze, message=FALSE}
analysis <- analyze_cat_multi(data = benchmarked_points,
                        weights = point_weights,
                        id_var = "unique_id",
                        cat_var = "Condition.Category",
                        wgt_var = "WGT",
                        split_vars = "Indicator",
                        definitions = cat_defs,
                        conf = confidence,
                        verbose = FALSE)
```

### Confidence Intervals
```{r confidence intervals, message=FALSE}
# Calculate CIs for acres
total_acres = (analysis$total_observation_weight/analysis$weighted_observation_proportion)[1]

analysis <- analysis %>% 
  mutate(lower_bound_acres = weighted_observation_proportion_lower_bound_80pct*total_acres,
         upper_bound_acres = weighted_observation_proportion_upper_bound_80pct*total_acres)

analysis$Reporting.Unit <- output_filename

names(analysis) <- c("Rating",
                     "Number of plots",
                     "Unweighted proportion of sampled area",
                     "Estimated Area",
                     "Weighted proportion of sampled area",
                     paste0(c("Lower confidence bound of percent area (", "Upper confidence bound of percent area ("), confidence, "%, Goodman multinomial) "),
                     "Indicator",
                     paste0(c("Lower confidence bound of acres (", "Upper confidence bound of acres ("), confidence, "%, Goodman multinomial)"),
                     "Reporting Unit")

```

### Formatting Outputs
```{r formatting outputs, message=FALSE}
wgtcat_summary <- weight_info[["frame.summary"]]
names(wgtcat_summary) <- c("wgtcat",
                           "total_count",
                           "count_Target Sampled",
                           "count_NT",
                           "count_IA",
                           "count_NA",
                           "count_UNK",
                           "area_ha",
                           "observed_proportion",
                           "observed_area",
                           "weight")

fates <- c("NT","IA","NA","UNK")

for (fate in fates) {
  wgtcat_summary[[paste0("count_", fate)]][is.na(wgtcat_summary[[paste0("count_", fate)]])] <- 0
}

wgtcat_summary[["in_inference"]] <- (wgtcat_summary[["weight"]])>0

wgtcat_summary$area_units <- "hectares"

wgtcat_summary$unobserved_area <- wgtcat_summary$area_ha - wgtcat_summary$observed_area

# Add in where they're from
all_points <- sp::merge(x = all_points,
                        y = point_weights[,c("unique_id","WGT")],
                        all.x = TRUE,
                        by = "unique_id")

# reformat benchmark_points to wide format
wgtcat_spdf <- merge(x = wgtcat_spdf,
                     y = wgtcat_summary,
                     by = "wgtcat")

# Rounding numbers for clarity
wgtcat_spdf@data[,is.numeric(wgtcat_spdf@data)] <- round(is.numeric(wgtcat_spdf@data), 2)
```
## Writing Outputs
```{r writing outputs, message=FALSE, warning=FALSE}
write.csv(point_weights,
          file = paste0(output_path, "/",
                        output_filename,
                        "_pointweights_",
                        date, ".csv"),
          row.names = FALSE)
write.csv(wgtcat_summary,
          file = paste0(output_path, "/",
                        output_filename,
                        "_wgtcatsummary_",
                        date, ".csv"),
          row.names = FALSE)
write.csv(analysis,
          file = paste0(output_path, "/",
                        output_filename,
                        "_analysis_",
                        date, ".csv"),
          row.names = FALSE)
rgdal::writeOGR(all_points,
                dsn = output_path,
                layer = paste0(output_filename,
                               "_points_",
                               date),
                driver = "ESRI Shapefile",
                overwrite_layer = TRUE)

rgdal::writeOGR(wgtcat_spdf,
                dsn = output_path,
                layer = paste0(output_filename,
                               "_wgtcats_",
                               date),
                driver = "ESRI Shapefile",
                overwrite_layer = TRUE)

# lets also write these to gdb to avoid horrible truncated field names...
arc.write(data = all_points,
          path = paste0(output_path,"/",
                        analysis_gdb,"/",output_filename,
                               "_points"),
          overwrite = TRUE)

arc.write(data = wgtcat_spdf,
          path = paste0(output_path,"/",
                        analysis_gdb,"/",output_filename,
                               "_wgtcats"),
          overwrite = TRUE)
```
## Results
### Benchmarks
```{r, message=FALSE}
benchmarks_table <- benchmarks %>% DT::datatable(extensions = 'Buttons', filter = "top" , 
                            options = list(scrollX = TRUE ,
                            dom = 'Bfrtip',
                            buttons =
                            list(list(
                            extend = 'collection',
                            buttons = c('csv', 'excel'),
                            text = 'Download Table'))) , 
                 caption = paste("Weighted Analysis Benchmarks for " , 
                                 toString(output_filename)) , 
                 rownames = FALSE)
benchmarks_table
```

### Point Weights
```{r, message=FALSE}
weights_table <- point_weights %>% DT::datatable(extensions = 'Buttons', filter = "top" , 
                            options = list(scrollX = TRUE ,
                            dom = 'Bfrtip',
                            buttons =
                            list(list(
                            extend = 'collection',
                            buttons = c('csv', 'excel'),
                            text = 'Download Table'))) , 
                 caption = paste("Point Weights for " , 
                                 toString(output_filename)) , 
                 rownames = FALSE)
weights_table
```
### Figures

#### Proportional Estimates Tables
```{r, fig.width= 12, fig.height=8, message=FALSE}
# use dt table to present analysis and wgtcat summary
library(DT)

# round numeric
analysis <- analysis %>%
  mutate_if(is.numeric, round, 2)

analysis_table <- analysis %>%
  DT::datatable(extensions = 'Buttons', filter = "top" , 
                            options = list(scrollX = TRUE ,
                            dom = 'Bfrtip',
                            buttons =
                            list(list(
                            extend = 'collection',
                            buttons = c('csv', 'excel'),
                            text = 'Download Table'))) , 
                 caption = paste("Weighted Analysis Results in " , 
                                 toString(output_filename)) , 
                 rownames = FALSE)

analysis_table

```

#### Weight Category Summaries
```{r, fig.width= 12, fig.height=8, message=FALSE, warning=FALSE}
wgtcat_table <- wgtcat_summary %>%
  DT::datatable(extensions = 'Buttons', filter = "top" , 
                            options = list(scrollX = TRUE ,
                            dom = 'Bfrtip',
                            buttons =
                            list(list(
                            extend = 'collection',
                            buttons = c('csv', 'excel'),
                            text = 'Download Table'))) , 
                 caption = paste("Weight Category Summary in  " , 
                                 toString(output_filename)) , 
                 rownames = FALSE)
wgtcat_table
```

#### Plotting weighted analysis results
```{r, fig.width=12, fig.height=8, message=FALSE}
# Fancy palette from Nelson
library(scales)
palette = c("#f5bb57ff",
                                            "#f8674cff",
                                            "#4a8b9fff",
                                            "#685b7fff",
                                            "#c95294ff",
                                            "#f5a9c6ff",
                                            "#75c6c5ff",
                                            "#fd6794ff")

analysis$Rating <- as.factor(analysis$Rating)

levels(analysis$Rating)

analysis$Rating <- factor(analysis$Rating,levels(factor(analysis$Rating))[c(2,3,4,1,5)])

# Join in nice looking name

analysis_merge <- merge(x = analysis,
                  y = unique(benchmarks[,c("Objective.name", "Management.Question")]),
                  by.x = "Indicator",
                  by.y = "Objective.name",
                  all.x = TRUE,
                  all.y = FALSE)

g <- ggplot(data = analysis_merge,
                aes(y = `Estimated Area`,
                             x = Rating, fill = Rating)) +
  geom_col(alpha = 0.6, position = "dodge") +
  geom_errorbar(aes(ymin = `Lower confidence bound of acres (80%, Goodman multinomial)`,
                                      ymax = `Upper confidence bound of acres (80%, Goodman multinomial)`),
                         width = 0.5)+
  facet_wrap(.~Management.Question, scales = "free")+
  coord_flip()+
  theme_bw()+
  labs(x = "")+
  scale_fill_manual(values = palette)+
  scale_y_continuous(label = comma)+
  theme(panel.spacing.y = unit(2,"lines"))

ggplotly(g)

ggsave(plot = g, device = "jpeg", dpi = 400, width =20, height = 10, filename = paste0("analysis_", output_filename, ".jpeg"), path = output_path)


```


